Parameter-less Population Pyramid

### 2014-01-13 Parameterless Population Pyramid/main.tex
Some wordsmithing

\maketitle
\begin{abstract}
%Maximum 200 Words
  A major limiting factor to practical application of evolutionary
  techniques is the tuning of their parameters which need to be
  reconfigured to each new problem.  While some previous methods been
  able to remove or reduce the need for parameter tuning, many do so
  by trading peak efficiency for general applicability.  The
  Parameter-less Population Pyramid (P3) is an evolutionary technique
  requiring \textbf{no} parameters and still broadly applicable to
  many problem types. It strikes a balance between continuous
  integration of diversity with exploitative elitist operators,
  allowing it to solve easy problems quickly and hard problems
  eventually.  When compared with three state of the art optimization
  techniques across seven different problem classes, P3 always finds
  the optimum at least a constant factor \emph{\textbf{faster}} than
  the rest.  Even when optimally tuning comparison technique
  parameters, P3 has a lower growth complexity when test problems come
  from a random distribution.  Beyond empirical results, we suggest
  avenues for theoretical analysis of P3's expected runtime.  In
  general, while these are only the first results for P3, we feel they
  are indicative wide applicability.
\end{abstract}

@@

Often built into the design of a genetic algorithm (GA) are
evolutionary parameters which allow the user to tune the optimizer's
assumptions concerning a specific problem.  These parameters can allow
a GA to more efficiently search the problem's landscapes.  However, to
do so the user must first determine how to correctly \textbf{tune} the
parameters, typically a complex and time consuming process unto
itself.

Some considerable work has been devoted to understanding how to tune
these parameters. Early work in Evolutionary Strategies introduced the
concept of self-adaptive evolution which was subsequently introduced
to genetic algorithms~\cite{Back:1992:selfadapt}. Here, each solution
carries some of the parameters used and those parameters are adapted
along with the solution they are co-encoded with. Other work introduced the concept
of a \emph{meta} algorithm\cite{grefenstette:1986:optimalga}. Here a
second optimizer runs ``on top'' of the actual problem-solving
optimizer, the second slowly tuning the parameters of the first. Both
of these approaches are rather slow, in particular the meta
approach, as actual problem solving must occur at set levels to give
feedback as to how to change the parameters (online or to the meta
optimizer).

Many of these parameters have ``reasonable'' settings that are
unlikely to effect the order of complexity of an genetic algorithm,
but population size is an exception. For example, a population sized
too small can cause exponential search times, while arbitrarily large
population sizes can lead to wasted evaluations and stalled search.
More importantly, setting the population size correctly relies on
difficult to discern features of each problem. Some research has been
done on setting optimal population
sizes. Goldberg~\cite{goldberg:1991:gasize} showed that small
population sizes were not capable of finding good solutions, though
some practitioners~\cite{haupt:2000:optimum} have found that small
populations with a large mutation rate are effective on practical
problems.

All of these complications have lead to research on how to rid
evolutionary/genetic algorithms of parameter settings of any kind.  In
an early example, a crossover based GA was devised which used
concurrent racing populations to remove all user specified
parameters~\cite{harik:1999:parameterlessga}.  This method of
population racing has also been paired with more advanced evolutionary
system, such as Hierarchical Bayesian
Optimization~\cite{pelikan:2004:parameterlesshboa}. The claim was that
while not optimal, this method ensured the GA would take no more than
a logarithmic increase in evaluations.  Furthermore, doing so would
ensure that these optimizers could be generally applicable without
problem specific tuning.  Experimental evidence supports this theory,
with some evidence that in general the bound is closer to a constant
factor worse than optimal.

@@

\section{The P3 Algorithm}
\label{sec-p3}
Before going into detail about how each aspect of P3 works, let us
consider an example of P3 operation. Unlike many EC techniques, P3
does not have a single population of solutions.  Instead P3 maintains
a pyramid like structure, such that each level of the pyramid has a
population of solutions, with each level representing something akin
to different generations of evolution. Consider the example evolution
presented in Figure~\ref{fig-p3}.  At the start of optimization there
are no solutions so solution \#1 is randomly generated.  Solution \#1
is brought to a local optima using a hill climber
(Section~\ref{sec-hillclimber}) creating solution \#2, which is added
to the population pyramid (Section~\ref{sec-pyramid}) at population
level $P_0$.  With a population of 1 at $P_0$, little more can be done
so solution \#3 is randomly generated and optimized via hill climbing
to create solution \#4, which is likewise added to the population
pyrimad at population level $P_0$.  With a population of two,
crossover (Section~\ref{sec-crossover}) is applied to solutions \#2
and \#4 creating solution \#5, which is an improvement. As \#5 used
information from $P_0$ it is added to the next level of the pyramid at
population level $P_1$.  Because of the crossover technique that P3
uses, there is no more information to be gained from the present
population. Thus, \#6 is randomly
generated and improved into \#7.  Unlike \#2 and \#4, \#7 is not added
in this example because it happens to be identical to a solution
already in the pyramid.  Even though \#7 is not unique, it still
crosses with $P_0$, which does make a unique improvement \#8 that is
added to $P_1$ and mixed with that level to create \#9.  As before,
the process creates \#10, \#11, and \#12 using random generation, hill
climbing, and crossover, respectively.  In this example \#13 is not
added to $P_2$ because it not a strict improvement in fitness over
\#12.  There is not enough information in $P_2$ to improve \#13
further, so the process would begin yet again with a random solution.

@@

For our purposes, we use a First Improvement Hill Climber
(FIHC)\footnote{We use the name FHIC for an algorithm that likely has
  various names in the literature, but use it here with a descriptive
  name for clarity.}, a method of exploiting local information through
a series of modify-evaluate-accept iterations starting from a randomly
initialized solution and ending with a optimum, either local or global.
FIHC uses the following algorithm.  Consider all gene loci
$x_i$ in a random order.  For a given $x_i$ try all possible
alternative values for the gene $x_i$, evaluating solution quality
with each change.  If any alternative value is a strict fitness improvement
over the current value for $x_i$, replace $x_i$ with the improved
value.  Iteration of the process continues until no such change results in
a strict fitness improvement.

@@

\subsection{Crossover}
\label{sec-crossover}
P3 uses a crossover method derived from that defined for LTGA~\cite{thierens:2013:ltgahiff}
(Section~\ref{sec-ltga}),
P3 uses gene value entropy to construct clusters of genes that should have their values
conserved during crossover.  There are two primary steps in
defining this type of crossover: \emph{cluster creation} and \emph{cluster usage}.

@@

Not all of  the resulting clusters in the linkage  tree are useful for
crossover.  The most obvious useless  cluster is one that contains all
the loci.   A crossover that preserves  all loci in  a solution cannot
create  a  new  individual.   We  further  remove  clusters  that  are
completely  contained   within  some  other  cluster,   such  that  if
$D(C_i,C_j)  =  0$, $C_i$  and  $C_j$  are  not used  when  performing
crossover.   $D(C_i,C_j)$ only  returns  $0$ when  the  loci in  $C_i$
perfectly predict the  bits in $C_j$ and vice versa. From this we can
conclude  that  the  settings  for  those two  loci  should  never  be
disrupted.  When  used on particularly small populations  this has the
potential  to  greatly  reduce  the  number of  clusters  used  during
crossover.  For  instance, if  $|P_i| = 1$,  no clusters exist. If
$|P_i| =  2$ at most two clusters  exist: one for all  genes where the
solutions agree and one for where they disagree.

@@

To complete a single crossover event, all of the useful clusters are tested, each
requiring at most a single evaluation.  Clusters are tested in smallest first order,
with clusters of equal size ordered randomly.  This ensures smaller building blocks
are optimized before large changes are made to the genome, thus helping preserve
diversity.  This also helps prevent genetic hitchhikers, which make interfere with
entropy calculations.  As cluster building and application ordering are not dependent on
the position of loci in the genome, there is no positional bias in the search points
evaluated.

\subsection{All Together}
\label{sec-alltogether}
The pieces of P3 were specifically chosen to provide important capabilities.  First,
the hill climber provides an efficient tool for discovering pairwise linkage, as
search only ends when no single bit change can be made which improves fitness.
Therefore crossover, which relies on pairwise linkage, has high quality information
to leverage when determining clusters.

@@

\subsection{Contrasted With The Generational Model}
A major advantage of P3's structure is that it avoids the need to set
a fixed population size.  In a generational model, the diversity
contained in the initial population represents a bottleneck for future
generations.  This is true because in most systems, especially on
deceptive problems, mutation rates are low enough as to make the
creation of large, multi-gene substructures unlikely after
initialization.  As such the generational model often results in a
race between decreasing diversity and increasing fitness. As such, the
population size must be set large enough to contain sufficient initial
diversity and to maintain that diversity during search for just long
enough for all of it to be leveraged.

@@

As P3 has exploited all known information each time the top of the pyramid is reached,
no evaluations are wasted  on unused diversity.  Similarly, because P3
does not throw away previously optimized solutions while adding diversity, it avoids
waste caused by premature convergence. It is this essential balance of
evaluation to diverstiy that P3 focuses on.

@@

To cause an update to $P_i$, all of the crossovers discovered for
$P_{i-1}$ must first be applied.  Each causes their own evaluation.
If P3 were to na\"ively use all crossovers and perform an evaluation
each time, this would cause $O(N)$ evaluations.  Again, this would
result in $O(N)$ amortized cost per update to $P_i$.  However,
Section~\ref{sec-crossover} discussed how P3 will ignore some clusters
and will also skip an evaluation if no change to the solution was made
by a crossover.  As such we currently are unable to state conclusively
that updating $P_i$ will require $O(N)$ evaluations.  That said,
empirically it appears to be true.  Furthermore, it seems counterintuitive
that intentionally wasting evaluations would be considered beneficial
to runtime.

@@

TODO Consider discussing memory usage.

@@

\subsection{Random Restart First Improvement Hill Climber}
TODO Need a citation either here or Section~\ref{sec-hillclimber}.

@@

If P3's quality relies solely on starting at random points and
applying FIHC, this comparison algorithm should perform even better
than P3.  Like P3, FHIC-alone algorithm is Parameter-less, makes no
assumptions about gene ordering, and will find the global optima given
a sufficient (potentially exponential) number of evaluations.

@@

Since LTGA requires a population size parameter, we thought it most fair to
compare against LTGA with an optimized population size, even though P3
requires no such tuning. To set the population size parameter, we use
the bisection method~\cite{goldman:2012:ltga}, which determines the minimum population
size required to meet a specified success criteria.  As P3 will run until the global optima
is found or the evaluation limit is reached, we want to ensure LTGA has a similar requirement
on its success.  Therefore, using the Rule of Three~\cite{jovanovic:1997:ruleofthree}, we
can provide a bound on the probability of failure $\frac{3}{k+1}$ with $95\%$ confidence,
where $k$ is a number of runs without a failure.  Here we state that a population size
is successful if it performs 100 runs without failing to find the global optimum, meaning
the probability of that population failing to solve problems of the same class is bounded above by
$3\%$.  Note that these 100 runs are considered training, and the results from the bisection
runs are not included during the comparative testing.  Furthermore, when testing on problem
classes, bisection is performed on a different set of instances than testing.

@@

Our final problem class is a subset of the Maximum Satisfiability (MAXSAT) problem.
This problem consists of a series of disjunctive clauses, such that each clause contains
three problem variables, some of which are negated. While
MAXSAT requires exponential time to solve, we have devised a method of generating
instances that ensures each has at least one solution that satisfies all clauses.
First, a random target solution is generated.  For each clause, set the sign of each
variable in the clause such that that clause will be satisfied by the target solution.
To ensure there is no bias to variable signs in the generated problem there is a $\frac{1}{6}$ probability
all clause signs match the target, a $\frac{1}{6}$ probability two of the three signs match,
and a $\frac{4}{6}$ probability only one sign matches.  By setting the clause signs
in this way we ensure the target solution is always able to satisfy all clauses,
meaning all problems have a known best fitness.  Note that its possible there to be
other solutions that also satisfy all clauses.  We chose a clause to variable ratio
of $r=4.27$.

@@

In total, our experiments performed 100 runs of each optimization algorithm across
7 benchmark problems, with each run limited to 100 million evaluations.
In performing data collection, we performed over 150,000 runs, totaling approximately
1.5 trillion evaluations.  Of these, approximately 100,000 runs were used to optimally
set the population size of LTGA.  Note that these runs should be considered a training phase,
and as such all reported results for LTGA use independent runs to create a testing phase.
This was the only algorithm tested that received such tuning.  As such, LTGA should have a significant
advantage over the others in that it has pre-learned the correct population size.
All of the code used to perform our experiments
as well as individual run results and analysis scripts are available from our
website\footnote{\url{https://name.of.site.removed/anonymous}}.  The results
are summarized in Figure~\ref{fig-results} and Table~\ref{table-results}.

@@

Using relaxed requirements, this can likely be extended to any arbitrary fully decomposable
problem. Assuming that for sufficiently large population sizes the entropy between
bits in a single subproblem is lower than the entropy of bits between subproblems,
P3 will construct a cluster capable of moving exactly one subproblem's gene values
during crossover.  As crossover cannot decrease the fitness of a solution each
crossover will have a non zero probability of improving subproblem solution quality.
The probability of low fitness local optima for a subproblem reaching high levels
is therefore low, which increases the probability of donors only having high quality
solutions to subproblems.  Combined this is expected to create a runtime potentially
exponential in subproblem size, but polynomial in problem size.

@@

Currently, P3 is limited to single objective, noiseless, binary encoded problems.
In future work we plan to extend (or show no modification is necessary) P3 to handle
problems from these classes.  Doing so will help make P3 more applicable to a wider
range of real world problems.

>>>>

\maketitle
\begin{abstract}
%Maximum 200 Words
  Real world applications of evolutionary techniques are often hindered by
  the need to determine problem specific parameter settings.
  While some previous methods have reduced or removed
  the need for parameter tuning, many do so
  by trading peak efficiency for general applicability.  The
  Parameter-less Population Pyramid (P3) is an evolutionary technique
  requiring \textbf{no} parameters which is still broadly applicable to
  many problem types. P3 strikes a balance between continuous
  integration of diversity with exploitative elitist operators,
  allowing it to solve easy problems quickly and hard problems
  eventually.  When compared with three state of the art optimization
  techniques across seven different problem classes, P3 always finds
  the optimum at least a constant factor \emph{\textbf{faster}} than
  the rest.  Even when optimally tuning comparison technique
  parameters, P3 has a lower evaluations to success growth complexity when test problems come
  from a random distribution.  Beyond empirical results, we suggest
  avenues for theoretical analysis of P3's expected runtime.  In
  general, while these are only the first results for P3, we feel they
  indicate wide applicability.
\end{abstract}

@@

Often built into the design of a genetic algorithm (GA) are
evolutionary parameters which allow the user to tune the optimizer's
assumptions to match a specific problem.  These parameters can allow
a GA to more efficiently search the problem's landscape.  However, to
do so the user must first determine how to correctly \textbf{tune} the
parameters, typically a complex and time consuming process unto
itself.

Some considerable work has been devoted to understanding how to tune
these parameters. Early work in Evolutionary Strategies introduced the
concept of self-adaptive evolution which was subsequently introduced
to genetic algorithms~\cite{Back:1992:selfadapt}. Here, each solution
carries some of the parameters used and those parameters are adapted
along with the solution they are co-encoded with. Other work introduced the concept
of a \emph{meta} algorithm~\cite{grefenstette:1986:optimalga}. Here a
second optimizer runs ``on top'' of the actual problem-solving
optimizer, the second slowly tuning the parameters of the first. Both
of these approaches are rather slow, in particular the meta
approach, as actual problem solving must occur at set levels to give
feedback as to how to change the parameters (online or to the meta
optimizer).

Many parameters have ``reasonable'' settings that are
unlikely to effect the order of complexity of a genetic algorithm,
with population size being a notable exception. For example, an overly small
population can cause exponential search times, while arbitrarily large
population sizes can lead to wasted evaluations and stalled search.
More importantly, setting the population size correctly relies on
difficult to discern features of each problem. 
Goldberg~\cite{goldberg:1991:gasize} showed that small
population sizes were not capable of finding good solutions, though
some practitioners~\cite{haupt:2000:optimum} have found that small
populations with a large mutation rate are effective on some practical
problems.

These complications have encouraged work to remove the need for
operator defined evolutionary parameters. In
an early example, a crossover based GA was devised which used
concurrent racing populations to remove all user specified
parameters~\cite{harik:1999:parameterlessga}.  This method of
population racing has also been paired with more advanced evolutionary
system, such as Hierarchical Bayesian
Optimization~\cite{pelikan:2004:parameterlesshboa}. The claim was that
while not optimal, this method ensured the GA would take no more than
a logarithmic factor increase in evaluations beyond optimal.  Furthermore, doing so would
ensure that these optimizers could be generally applicable without
problem specific tuning.  Experimental evidence supports this theory,
with some evidence that in general the bound is closer to a constant
factor worse than optimal.

@@

\section{The P3 Algorithm}
\label{sec-p3}
Before going into detail about how each aspect of P3 works, let us
consider an example of P3 operation. Unlike many EC techniques, P3
does not have a single population of solutions.  Instead P3 maintains
a pyramid like structure, such that each level of the pyramid has a
population of solutions, with each level representing something akin
to different generations of evolution. Consider the example evolution
presented in Figure~\ref{fig-p3}.  At the start of optimization P3
has no information about the landscape so solution \#1 is randomly generated.  Solution \#1
is brought to a local optima using a hill climber
(Section~\ref{sec-hillclimber}) creating solution \#2, which is added
to the population pyramid (Section~\ref{sec-pyramid}) at population
level $P_0$.  With a population of 1 at $P_0$, there is insufficient information to continue,
so solution \#3 is randomly generated and optimized via hill climbing
to create solution \#4, which is likewise added to the population
pyrimad at population level $P_0$.  With a population of two,
crossover (Section~\ref{sec-crossover}) is applied to solutions \#2
and \#4 creating solution \#5, which is an improvement. As \#5 used
information from $P_0$ it is added to the next level of the pyramid at
population level $P_1$.  Because of the crossover technique that P3
uses, there is no more information to be gained from the present
population. Thus, \#6 is randomly
generated and improved into \#7.  Unlike \#2 and \#4, \#7 is not added
in this example because it happens to be identical to a solution
already in the pyramid.  Even though \#7 is not unique, it still
crosses with $P_0$, which does make a unique improvement \#8 that is
added to $P_1$ and mixed with that level to create \#9.  As before,
the process creates \#10, \#11, and \#12 using random generation, hill
climbing, and crossover, respectively.  In this example \#13 is not
added to $P_2$ because it not a strict fitness improvement over
\#12.  There is not enough information in $P_2$ to improve \#13
further, so the process would begin yet again with a random solution.

@@

For our purposes, we use a First Improvement Hill Climber
(FIHC)\footnote{We use the name FHIC for an algorithm that likely has
  various names in the literature, but use it here with a descriptive
  name for clarity.}, a method of exploiting local information through
a series of modify-evaluate-accept iterations starting from a randomly
initialized solution and ending with an optimum, either local or global.
FIHC uses the following algorithm.  Consider all gene loci
$x_i$ in a random order.  For a given $x_i$ try all possible
alternative values for the gene $x_i$, evaluating solution quality
with each change.  If any alternative value is a strict fitness improvement
over the current value for $x_i$, replace $x_i$ with the improved
value.  Iteration of the process continues until no such change results in
a strict fitness improvement.

@@

\subsection{Crossover}
\label{sec-crossover}
P3 uses a crossover method derived from that defined for LTGA~\cite{thierens:2013:ltgahiff}
(Section~\ref{sec-ltga});
P3 uses gene value entropy to construct clusters of genes that should have their values
conserved during crossover.  There are two primary steps in
defining this type of crossover: \emph{cluster creation} and \emph{cluster usage}.

@@

Not all of  the resulting clusters in the linkage  tree are useful for
crossover.  The most obvious useless  cluster is one that contains all
the loci.   A crossover that preserves  all loci in  a solution cannot
create  a  new  individual.   We  further  remove  clusters  that  are
completely  contained   within  some  other  cluster,   such  that  if
$D(C_i,C_j)  =  0$, $C_i$  and  $C_j$  are  not used  when  performing
crossover.   $D(C_i,C_j)$ only  returns  $0$ when  the  loci in  $C_i$
perfectly predict the gene values in $C_j$ and vice versa. From this we can
conclude  that  the  settings  for  those two  loci  should  never  be
disrupted.  When  used on particularly small populations  this has the
potential  to  greatly  reduce  the  number of  clusters  used  during
crossover.  For  instance, if  $|P_i| = 1$,  no clusters exist. If
$|P_i| =  2$ at most two clusters  exist: one for all  genes where the
solutions agree and one for where they disagree.

@@

To complete a single crossover event, all of the useful clusters are tested, each
requiring at most a single evaluation.  Clusters are tested in smallest first order,
with clusters of equal size ordered randomly.  This ensures smaller building blocks
are optimized before large changes are made to the genome, thus helping preserve
diversity.  This also helps prevent genetic hitchhikers, which may interfere with
entropy calculations.  As cluster building and application ordering are not dependent on
the position of loci in the genome, this operator introduces no positional bias in the search points
evaluated.

\subsection{All Together}
\label{sec-alltogether}
The pieces of P3 were specifically chosen to provide important complimentary capabilities.  First,
the hill climber provides an efficient tool for discovering pairwise linkage, as
search only ends when no single bit change can be made which improves fitness.
Therefore crossover, which relies on pairwise linkage, has high quality information
to leverage when determining clusters.

@@

\subsection{Contrasted With The Generational Model}
A major advantage of P3's structure is that it avoids the need to set
a fixed population size.  In a generational model, the diversity
contained in the initial population represents a bottleneck for future
generations.  This is true because in most systems, especially on
deceptive problems, mutation rates are low enough as to make the
creation of large, multi-gene substructures unlikely after
initialization.  As such the generational model often results in a
race between decreasing diversity and increasing fitness. This means the
population size must be set large enough to contain sufficient initial
diversity and to maintain that diversity during search for just long
enough for all of it to be leveraged.

@@

As P3 has exploited all known information each time the top of the pyramid is reached,
no evaluations are wasted  on unused diversity.  Similarly, because P3
does not throw away previously optimized solutions while adding diversity, it avoids
waste caused by premature convergence. It is this essential balance of
exploitation to diversity that P3 focuses on.

@@

To cause an update to $P_i$, all of the crossovers discovered for
$P_{i-1}$ must first be applied.  Each causes their own evaluation.
If P3 were to na\"ively use all crossovers and perform an evaluation
each time, this would cause $O(N)$ evaluations.  Again, this would
result in $O(N)$ amortized cost per update to $P_i$.  However,
Section~\ref{sec-crossover} discussed how P3 will ignore some clusters
and will also skip an evaluation if no change to the solution was made
by a crossover.  As such we are currently unable to state conclusively
that updating $P_i$ will require $O(N)$ evaluations.  That said,
empirically it appears to be true.  Furthermore, it seems counterintuitive
that intentionally wasting evaluations would be considered beneficial
to runtime.

@@

%TODO Consider discussing memory usage.

@@

\subsection{Random Restart First Improvement Hill Climber}
%TODO Need a citation either here or Section~\ref{sec-hillclimber}.

@@

If P3's quality relies solely on starting at random points and
applying FIHC, this comparison algorithm should perform even better
than P3.  Like P3, the FHIC-alone algorithm is parameter-less, makes no
assumptions about gene ordering, and will find the global optima given
a sufficient (potentially exponential) number of evaluations.

@@

Since LTGA requires a population size parameter, we thought it most fair to
compare against LTGA with an optimized population size, even though P3
receives no such tuning. To set the population size parameter, we use
the bisection method~\cite{goldman:2012:ltga}, which determines the minimum population
size required to meet a specified success criteria.  As P3 will run until the global optima
is found or the evaluation limit is reached, we want to ensure LTGA has a similar requirement
on its success.  Therefore, using the Rule of Three~\cite{jovanovic:1997:ruleofthree}, we
can provide a bound on the probability of failure $\frac{3}{k+1}$ with $95\%$ confidence,
where $k$ is a number of runs without a failure.  Here we state that a population size
is successful if it performs 100 runs without failing to find the global optimum, meaning
the probability of that population failing to solve problems of the same class is bounded above by
$3\%$.  Note that these 100 runs are considered training, and the results from the bisection
runs are not included during the comparative testing.  Furthermore, when testing on problem
classes, bisection is performed on a different set of instances than testing.

@@

Our final problem class is a subset of the Maximum Satisfiability (MAX-SAT) problem.
This problem consists of a series of disjunctive clauses, such that each clause contains
three problem variables, some of which are negated. While
MAX-SAT requires exponential time to solve, we have devised a method of generating
instances that ensures each has at least one solution that satisfies all clauses.
First, a random target solution is generated.  For each clause, set the sign of each
variable in the clause such that that clause will be satisfied by the target solution.
To ensure there is no bias to variable signs in the generated problem there is a $\frac{1}{6}$ probability
all clause signs match the target, a $\frac{1}{6}$ probability two of the three signs match,
and a $\frac{4}{6}$ probability only one sign matches.  By setting the clause signs
in this way we ensure the target solution is always able to satisfy all clauses,
meaning all problems have a known best fitness.  Note that its possible there to be
other solutions that also satisfy all clauses.  We chose a clause to variable ratio
of $r=4.27$.

@@

In total, our experiments performed 100 runs of each optimization algorithm across
7 benchmark problems, with each run limited to 100 million evaluations.
In performing data collection, we performed over 150,000 runs, totaling approximately
1.5 trillion evaluations.  Of these, approximately 100,000 runs were used to optimally
set the population size of LTGA.  Note that these runs should be considered a training phase,
and as such all reported results for LTGA use independent runs to create a testing phase.
This was the only algorithm tested that received such tuning.  As such, LTGA should have a significant
advantage over the others in that it has pre-learned the correct population size.
All of the code used to perform our experiments
as well as individual run results and analysis scripts are available from our
website.\footnote{\url{https://name.of.site.removed/anonymous}}  The results
are summarized in Figure~\ref{fig-results} and Table~\ref{table-results}.

@@

Using relaxed requirements, this can likely be extended to any arbitrary fully decomposable
problem. Assuming that for sufficiently large population sizes the entropy between
bits in a single subproblem is lower than the entropy of bits between subproblems,
P3 will construct a cluster capable of moving exactly one subproblem's gene values
during crossover.  As crossover cannot decrease the fitness of a solution each
crossover will have a greater than zero probability of improving subproblem solution quality.
The probability of low fitness local optima for a subproblem reaching high pyramid levels
is therefore low, which increases the probability of donors only having high quality
solutions to subproblems.  Combined this is expected to create a runtime potentially
exponential in subproblem size, but polynomial in problem size.

@@

Currently, P3 is limited to single objective, noiseless, binary encoded problems.
In future work we plan to extend (or show no modification is necessary for) P3 to handle
problems from these classes.  Doing so will help make P3 more applicable to a wider
range of real world problems.
